### 2020-12-26 examples.tex
fixed tings

% !TeX root = ./main.tex

\section{Examples}

\begin{example}
$\lambda_1 = (1,0,0)$, $\lambda_2 = (1,1,0)$, $\mu_1 = (0,1,0)$, $\mu_2 = (1,0,1)$. \jcom{$\mu = \mu_1 + \mu_2$ determines the blocks we have on the RHS of the BD MVy isomorphism of Equation~\ref{eq:topmvy2}.}

In the non-BD case, MVy establish
$$
\overline{\Gr^\lambda}\cap \cW^\mu \to 
\left\{ X = 
\left[\begin{BMAT}{cc|cc|c}{cc|cc|c}
0 & 1 & 0 & 0 & 0 \\
* & * & * & * & * \\
0 & 0 & 0 & 1 & 0 \\
* & * & * & * & * \\
* & 0 & * & 0 & * 
\end{BMAT}\right] \big| X \in \overline{\OO}_\lambda
\right\}
$$

In the BD case the RHS will consist of the same block like matrices $X$ but now having eigenvalues $s, 0$ such that $X - s\big|_{E_s}\in \OO_{\lambda_2}$ 
\end{example}

\begin{example}
Do Joel's exercise: It would be good to do an example where we will see some multiplicity in the fusion product.  I think that the simplest example would be the fusion product of the MV cycles for $\SL_3$ of weights $2\alpha_1$ and $2\alpha_2$.  Following the notation from the mvbasis paper, this would correspond to the following multiplication:
$x^2 y^2 = (xy-z)^2 + 2(xy - z) z + z^2$. 
For this example, I think we need $\lambda_1 = (2,0,0)$, $\lambda_2 = (2,2,0)$, $\mu_1 = (0,2,0)$, $\mu_2 = (2,0,2)$. 
\end{example}

\begin{example}
    Let 
    \[
    \begin{aligned}
        \mu_1 &= (2) & \lambda_1 &= & \mu &= (5) \\
        \mu_2 &= (3) & \lambda_2 &= & \lambda &= 
    \end{aligned}
    \]
    Consider the companion matrix $C(p)$ of 
    $$p(t) = (t-s)^3t^2 = (t^3 - 3t^2 s + 3t s^2 - s^3)t^2 = t^5 - 3 t^4 s + 3t^3 s^2 - t^2 s^3$$ 
    Let $X = C(p)^T$ so 
    \[
    X = \left[\begin{BMAT}(e){ccccc}{ccccc}
        0 & 1 & & & \\ 
        & 0 & 1 & & \\
        &   & 0 & 1 & \\ 
        &   &   & 0 & 1 \\
       0 & 0 & s^3 & -3s^2  & 3s  
    \end{BMAT}\right]    
    \]
    Ask that $X\big|_{E_0}$ has Jordan type $\lambda_1$ and $X-s\big|_{E_s}$ has Jordan type $\lambda_2$. In this rank 1 case we are forced to take $\lambda_i = \mu_i$.

    So what are the generalized eigenspaces $E_i$ ($i = 1,2$)? Note $\dim E_0 = 2$ and $\dim E_s = 3$.

    \acom{The basis
    \[
     [1], [t], [\bf 1], [\bf t], [\bf t^2]    
    \] 
    with $t[t] = 0$ and $t[{\bf t^2}] =   s^3 [{\bf 1}] - 3s^2 [{\bf t}] + 3s[{\bf t^2}]$ \textit{is not the correct basis to consider}.
    Hence my confusion of yore: what we would like is $t[t] = 0$ no? what the matrix is telling us is that $t[t] = [{\bf 1}]$. Can we still speak of \textit{two} generalized eigenspaces?}  % [3st^2 - 3s^2 t + s^3]

    Rather, take $B$ to be the basis $b_1 = e_1$, $b_2 = e_2$, $b_5 = e_5$, $b_4 = Xe_5$, $b_3 = X^2e_5)$. In this basis
    \[
    X_B = [X(b_i)] = \begin{bmatrix}
        0 & 1 & 0 & 0 & 0 \\
        0 & 0 & 0 & 0 & 0 \\
        0 & 0 & s & 1 & 0 \\
        0 & 0 & 0 & s & 1 \\
        0 & 0 & 0 & 0 & s
    \end{bmatrix}
    \]

    % I think this demonstrates why Roger's idea is probably the right thing to do. Namely, consider instead $T_{\mu_1,\mu_2}$ whose elements are 
    % \[
    %     X = \left[\begin{BMAT}(e){cc|ccc}{cc|ccc}
    %         0 & 1 & & & \\ 
    %         & 0 & x & y & z \\
    %         &   & 0 & 1 & \\ 
    %         &   &   & 0 & 1 \\
    %        0 & 0 & s^3 & -3s^2  & 3s  
    %     \end{BMAT}\right]    
    % \] 
    % such that $T_{\mu_1,\mu_2}\big|_{s = 0} = T_\mu$ whatever that means... 
\end{example}

\begin{example}
    Let 
    \[
    \begin{aligned}
        \mu_1 &= (3,1,1) & \lambda_1 &= (3,2,0) & \mu = (3,3,1) \\
        \mu_2 &= (0,2,0) & \lambda_2 &= (2,0,0) & \lambda = (5,2,0)
    \end{aligned}    
    \]
    and consider the companion matrices of 
    $$
    \begin{aligned}
        p_1(t) &= t^3 & p_2(t) &= t(t-s)^2 = t^3 - 2st^2 + s^2 t & p_3(t) &= t
    \end{aligned} % = t(t^2 - 2st + s^2) 
    $$
    \[
    X = \left[
        \begin{BMAT}(e){ccc;ccc;c}{ccc;ccc;c}
            0 & 1 & 0 & 0 & 0 & 0 & 0 \\
            0 & 0 & 1 & 0 & 0 & 0 & 0 \\
            % a & b & c & d & e & f & g \\ 
            0 & 0 & 0 & d & e & f & g \\ 
            0 & 0 & 0 & 0 & 1 & 0 & 0 \\
            0 & 0 & 0 & 0 & 0 & 1 & 0 \\
            0 & 0 & 0 & 0 & -s^2 & 2s & k \\
            0 & 0 & 0 & 0 & 0 & 0 & 0
        \end{BMAT}
        \right]    
    \]
\end{example}

\begin{example}
    Let
    \[
    \begin{aligned}
        \lambda_1 &= (3,2,0) & \mu_1 &= (3,1,1) \\
        \lambda_2 &= (2,0,0) & \mu_2 &= (1,1,0)
    \end{aligned}
    \]
    so the first MV cycle $Z_1 \cong \PP^1$ has MV polytope $\Conv\{0, \alpha_1\}$ and the second MV cycle $Z_2 \cong \PP^1$ has MV polytope $\Conv\{0, \alpha_2\}$. Their fusion product corresponds to two $\PP^2$'s intersecting along a $\PP^1$. That is, we have the fusion product
    \[
    Z_1 * Z_2 = Z_+ + Z_-
    \]
    where $Z_+ \cong Z_- \cong \PP^2$.
    We have 
    \[
    X = \begin{bmatrix}
    0 & 1 \\
      & 0 & 1 \\
      &   & 0 & 1 \\
      &   &   & s & a & b & c \\
      &   &   &   & 0 & 1 & 0 \\
      &   &   &   & 0 & s & d \\
      &   &   &   &   &   & 0
    \end{bmatrix}
    \]
    The $0$-generalized eigenspace $E_0$ of $X$ is $5$-dimensional, containing a 3-cycle and a 2-cycle.
    The 3-cycle is 
    \[
    \{X^2 e_3 , X e_3, e_3\} = \{e_1, e_2, e_3\}.
    \]
    To obtain another vector in $\ker X$, either $a = 0$ or $c = d = 0$, but the latter case cannot give a 2-cycle as $e_7 \notin \im X$. Then $a = 0$ and we obtain a 2-cycle
    \[
    \left\{X \left( e_6 -\frac{s}{d}e_7 \right), e_6 - \frac{s}{d}e_7 \right\}
    = \left\{e_5, e_6 - \frac{s}{d}e_7 \right\}.
    \]
    We also obtain the equations $b \neq 0$, $d \neq 0$, and $sc - bd = 0$ from this.

    For the $s$-generalized eigenspace $E_s$, we need $a + sb \neq 0$ to obtain a 2-cycle, which can be taken as 
    \[
    \begin{split}
    & \left\{ 
    (X -sI) \left( e_2 + 2s e_3 + 3s^2 e_4 + \frac{s^3}{a+sb} e_5 + \frac{s^4}{a+sb} e_6 \right), \right. \\ 
    & \quad \qquad \qquad  \left. e_2 + 2s e_3 + 3s^2 e_4 + \frac{s^3}{a+sb} e_5 + \frac{s^4}{a+sb} e_6 
    \right\} \\
    = & \left\{ 
    e_1 + s e_2 + s^2 e_3 + s^3 e_4, 
    e_2 + 2s e_3 + 3s^2 e_4 + \frac{s^3}{a+sb} e_5 + \frac{s^4}{a+sb} e_6
    \right\}
    \end{split}
    \]
    The minimal polymonial is $X^3 (X-sI)^2$, which when equated to 0 gives again the equation $cs-bd = 0$.
    Thus the defining equations are 
    \[
    \{ a = 0, cs-bd = 0\}.
    \]
    When we take $s = 0$, we get the equations
    \[
    \{a = 0, bd = 0\}
    \]
    which corresponds to two $\AA^2$'s intersecting along an $\AA^1$. This is indeed an open subset of $\PP^2 \cup_{\PP_1} \PP^2$, as required.
\end{example}

\begin{example}[Example 7 continued\dots]
    The matrix $X$ from the previous example defines (under MVy) the matrix 
    \[
        g = \begin{bmatrix}
            (t-s)t^3 \\
            -bt & (t-s)t \\
            -c & -d & t 
        \end{bmatrix}
    \]
    in $G(\cO)$. 
    Indeed the various blocks of $X$ are in a precise sense the companion matrices of the polynomial entries of $g$
    % First let's double check the inverse map.

    In $\Gr$ the element $g$ defines the lattice 
    \[
        gL_0 = \CC[t]\langle
        (t-s)t^3e_1 - (a+bt) e_2 - ce_3 , (t-s)te_2 - de_3, te_3 
        \rangle
    \]
    and computing the matrix of the action of $t$ on the quotient $L_0/L$ in the basis 
    \[
    \{
        [e_1],[te_1],[t^2e_1],[t^3e_1],[e_2],[te_2],[e_3]
    \}    
    \]
    recovers $X$ up to a transpose of course.

    Now let's see what we get when we invert $t$ and $t-s$ respectively.

    First let's invert $t$ by considering $L_2 = L\otimes\CC\xt[t-s]$. 
    \[
        L_2  = \CC[t,t^{-1}]\langle(t-s) e_1 -\frac{a+bt}{t^3} e_2 - \frac c {t^3} e_3, (t-s) e_2 - \frac d t e_3, e_3 \rangle
    \]
    so in $L_0/L_2$ we have 
    \[
    t[e_1] = s[e_1] + \frac{a+bt}{t^3} [e_2] + \frac c {t^3}  [e_3] \qquad t[e_2] = s[e_2] + \frac d t[e_3]    \qquad [e_3] = 0 
    \]
    and
    \[
    \left[t\big|_{L_0/L_2}\right]_{\{[e_1],[e_2]\}} = \begin{bmatrix}
        s \\
        \frac{a+bt}{t^3} & s 
    \end{bmatrix}
    \]
    which upon subtracting $s I$ gives a matrix having block type $\mu_2$ and Jordan type $\lambda_2 = (2)$ assuming $\frac{a+bt}{t^3}\ne 0$.

    Next let's invert $t-s$ by considering $L_1 = L\otimes\CC\xt$. 
    \[
    \begin{aligned}
        L_1 &= \CC[t,(t-s)^{-1}]\langle
        t^3 e_1 - \frac{a+bt}{t-s}e_2 - \frac c {t-s} e_3, te_2 - \frac d {t-s} e_3 , te_3 
        \rangle \\
        &= \langle
        t^3 e_1 - \frac{a}{t-s}e_2 - \frac b{t-s} te_2 - \frac c {t-s} e_3, te_2 - \frac d {t-s} e_3 , te_3 
        \rangle
    \end{aligned}
    \]
    so in $L_0/L_1$ we have 
    \begin{align*}
        t[e_1] &= [te_1] \\
        t[te_1] &= [t^2 e_1] \\ 
        t[t^2e_1] &= \frac{a}{t-s}[e_2] + \frac{b}{t-s} t[e_2] + \frac c {t-s} [e_3] \\
                  &= \frac b{t-s} \frac d{t-s} [e_3] + \frac c {t-s} [e_3] \\ 
                  &= \frac {bd + (t-s) c} {(t-s)^2} [e_3] \\
                  &= \frac{bd - sc}{(t-s)^2}[e_3] + \frac c {(t-s)^2} t[e_3] = 0 \\ 
        t[e_2] &= \frac d {t-s} [e_3] \\ 
        t[e_3] &= 0 
    \end{align*}
    and 
    \[
    \left[t\big|_{L_0/L_1}\right]_{\{[e_1],[te_1],[t^2e_1],[e_2],[e_3]\}} = \begin{bmatrix}
        0 \\
        1 & 0 \\
        0 & 1 & 0 \\
          & & 0 & 0 \\
          & & 0 & \frac d {t-s} & 0  
    \end{bmatrix}
    \]
    taking transpose 
    \[
        \begin{bmatrix}
            0 & 1\\
              & 0 & 1\\
              &   & 0 &  &  \\
              & & & 0 & \frac d {t-s} \\
              & & & & 0  
        \end{bmatrix}
    \]
    which has block type $\mu_1$ and Jordan type $\lambda_1 = (3,2)$ assuming $d\ne 0$.

    I have used the relations Roger found (and I checked) $a = 0$ and $cs - bd = 0$ above.

    To sum up the pair of matrices above should contain the same information as the matrix from the previous example 
    \[
    \begin{bmatrix}
    0 & 1 \\
      & 0 & 1 \\
      &   & 0 & 1 \\
      &   &   & s &  & b & c \\
      &   &   &   & 0 & 1 & 0 \\
      &   &   &   & 0 & s & d \\
      &   &   &   &   &   & 0
    \end{bmatrix} \Leftrightarrow \left( \begin{bmatrix}
            0 & 1\\
              & 0 & 1\\
              &   & 0 &  &  \\
              & & & 0 & \frac d {t-s} \\
              & & & & 0  
        \end{bmatrix} , \begin{bmatrix}
        s & \frac b {t^2} \\ & s 
        \end{bmatrix}\right) 
    \] 
\end{example}

{\bf Wish.} Given $L$ in $\Grbd$ define a map to $T_\mu$ just like MVy by taking $[t\big|_{L_0/L}]$ and use the fact that $[t\big|_{L_0/L_i]$ for $i = 1,2$ are companion matrices of the right type/piece together two MVy isomorphisms to make a BD MVy iso.

Lin alg question: if $p(t,t-s) = p_1(t)p_2(t-s)$ how are $C(p_1),C(p_2),C(p)$ related. I think it's basically this theorem \url{https://en.wikipedia.org/wiki/Structure_theorem_for_finitely_generated_modules_over_a_principal_ideal_domain}